### 2017-04-16 paper.tex
Initial

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2017}
\usepackage{times}
\usepackage{latexsym}

\usepackage{url}
\usepackage{booktabs}
\usepackage{graphicx}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{Toward a Comparable Corpora of Latvian, Russian and English Tweets}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
\end{abstract}

\citet{milajevs-griffiths:2016:RepEval}

\section{Introduction}
\label{sec:introduction}

\section{Dataset construction}
\label{sec:construction}

The initial set of tweets was retrieved using the \texttt{POST status/filter} endpoint of the Twitter Streaming API.\footnote{\url{https://dev.twitter.com/streaming/reference/post/statuses/filter}} The collected tweets had to be geo-located and had to originate from the area of Riga, the capital of Latvia.\footnotemark{}

\footnotetext{The \texttt{locations} parameter was set to \texttt{23.9325829, 56.8570671, 24.3247299, 57.0859184}}

251\,083 tweets were collected from the period from the 1st of November 2016 to the 31st of March 2017. On April 14th 2017, the collection was rehydrated by querying the Twitter API with the collected tweet IDs to get rid of the deleted tweets. In addition, the tweets that originated retweets were included to the collection: the JSON representation of a retweet includes the original tweet, which we extracted and added to the collection. Rehydrated and expanded collection resulted to 220\,883 tweets. Table~\ref{tab:tweet-counts} summarises the number of collected tweets.

\input{figures/tweet-counts.tex}

Further analysis of the extended rehydrated collection showed that there are 23115 (10.5\%) tweets that originated from the user check-ins with Swarm\footnote{\url{https://www.swarmapp.com/}} on Foursquare.\footnote{\url{https://foursquare.com/}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' most of the time follow a predefined template and thus do not reflect real language use.

\input{figures/client-counts.tex}

Table~\ref{tab:client-counts} shows the top ten most popular clients in the rehydrated collection. Together with the tweets originated from Foursquare, tweets from Instagram, an image sharing service), and Endomondo, a workout tracking service, were removed. Tweets written using the ``Word Cities'' client that posts weather reports and the Linkis client, a promotion website, were also removed.

The final collection resulted into 136\,067 tweets which are in Latvian, Russian or English and created after 1st of November 2016. The language of a tweet is provided by the corresponding field in the tweet JSON representation.

\input{figures/timeline.tex}

\section{Quantitative analysis}
\label{sec:quantitative}

Out of 136\,067 tweets that constitute the final collection, 45.5\% are in Latvian, 33.9\% are in Russian and 20.7\% are in English, see Table~\ref{tab:language-counts} for tweet counts. The ratio between the Latvian and Russian tweets is roughly the same as the proportion of ethnic Latvians and Russians in Riga, which is 46.2\% to 37.7\%.\footnote{\url{https://en.wikipedia.org/wiki/Riga}}

\input{figures/language-counts.tex}

Figure~\ref{fig:timeline} shows the bandwidth of tweets in time for all three languages. There are several peaks in Twitter usage.

\paragraph{Mid November}

18 November is the Proclamation Day of the Republic of Latvia. Numerous events take place across the country including torchlight processions and fireworks. The volume of tweets peaks in all three languages, thought the peak in Russian is less significant than in Latvian and English.

\paragraph{Mid December}

There is higher amount of Russian tweets, but Latvian and English tweet volume stays constant.

\paragraph{Early January}

There is a lot of activity during the New Year celebration for all three languages. Russian tweets peak the most almost reaching the same number of tweets as tweets in Latvian.

\paragraph{Late January}

In late January, the number of Latvian tweets increases, while for other languages it stays the same.

\paragraph{Mid March}

There is a general decline in Twitter usage for Latvian and English, but not for Russian, when for a short period of time there are more Russian tweets posted than Latvian.

\bibliography{references,dmilajevs_publications}
\bibliographystyle{acl_natbib}

\end{document}

### 2017-04-16 paper.tex
More on events

@@
 
-Figure~\ref{fig:timeline} shows the bandwidth of tweets in time for all three languages. There are several peaks in Twitter usage.
+Figure~\ref{fig:timeline} shows the bandwidth of tweets in time for all three languages.
+There are several peaks in Twitter usage.
+Some of them affect all three languages, as in early January, some of them affect only one language, as in late January.
+
+If the Twitter behaviour is affected by the events in the real world, then the peaks should correspond to events in the real word.
+The difference in peaks, could then be explained as there are different real word events that trigger discussions on Twitter in Latvian, Russian and English.
+Table~\ref{tab:timeline-lang-corr} suggests, that tweets in Latvian and English share similar behaviour.
+The Russian tweet timeline is distinct from both timelines, tough it's behaviour is more similar to the Latvian timeline than to English.
+
+\input{figures/timeline-lang-corr.tex}
+
+What are the distinctive and similar properties of the timelines? To answer the question, we first identify the real world events that happened during the biggest peaks.
 
@@
 
+11th of November is \href{https://en.wikipedia.org/wiki/L\%C4\%81\%C4\%8Dpl\%C4\%93sis_Day}{L\={a}\v{c}pl\={e}sis Day}, a memorial day for soldiers who fought for the independence of Latvia.
 18 November is the Proclamation Day of the Republic of Latvia.
-Numerous events take place across the country including torchlight processions and fireworks.
+Numerous events take place across the country including candle placing on and by the wall of Riga Castle, torchlight processions and fireworks.
 The volume of tweets peaks in all three languages, thought the peak in Russian is less significant than in Latvian and English.
 
-\paragraph{Mid December}
+% \paragraph{Mid December}
 
-There is higher amount of Russian tweets, but Latvian and English tweet volume stays constant.
+% There is higher amount of Russian tweets, but Latvian and English tweet volume stays constant.
 
@@
 
-In late January, the number of Latvian tweets increases, while for other languages it stays the same.
+The inauguration of Donald Trump as the 45th President of the United States held on 20th of January 2017. The number of Latvian tweets increases, while for other languages it stays roughly the same.
 
-\paragraph{Mid March}
+% \paragraph{Early March}
 
-There is a general decline in Twitter usage for Latvian and English, but not for Russian, when for a short period of time there are more Russian tweets posted than Latvian.
+% There are peaks in Latvian and Russian tweets.
+
+% \paragraph{Mid March}
+
+% There is a general decline in Twitter usage for Latvian and English, but not for Russian, when for a short period of time there are more Russian tweets posted than Latvian.
+
+Timeline analysis gives an insight of what events are reflected on Twitter, but does not indicate how twitter is used by ethnic communities, as language choice might depend on the topic, rather than author preference.
+
+\section{Language use}
+\label{sec:lang-use}
 

### 2017-04-16 paper.tex
Cite software.

@@
 
-\citet{milajevs-griffiths:2016:RepEval}
-
 \section{Introduction}
@@
 
-The initial set of tweets was retrieved using the \texttt{POST status/filter} endpoint of the Twitter Streaming API.\footnote{\url{https://dev.twitter.com/streaming/reference/post/statuses/filter}} The collected tweets had to be geo-located and had to originate from the area of Riga, the capital of Latvia.\footnotemark{}
+The initial set of tweets was retrieved by subscribing to the \texttt{POST status/filter} endpoint of the Twitter Streaming API\footnote{\url{https://dev.twitter.com/streaming/reference/post/statuses/filter}} with Poultry \cite{dmitrijs_milajevs_2017_546609}.
+The collected tweets had to be geo-located and had to originate from the area of Riga, the capital of Latvia.\footnotemark{}
 

### 2017-04-17 paper.tex
Language use

@@
 
+\input{figures/timeline.tex}
+
 The initial set of tweets was retrieved by subscribing to the \texttt{POST status/filter} endpoint of the Twitter Streaming API\footnote{\url{https://dev.twitter.com/streaming/reference/post/statuses/filter}} with Poultry \cite{dmitrijs_milajevs_2017_546609}.
@@
 
-\input{figures/timeline.tex}
-
-\section{Quantitative analysis}
-\label{sec:quantitative}
+\section{Timeline analysis}
+\label{sec:timeline}
 
@@
 
+\input{figures/score-hist.tex}
+
+We have seen that the reflection of an event on Twitter depends on the language.
+How are languages used individually?
+Do tweet authors tweet in a one language all the time, or switch between languages depending on the context?
+
+We consider 507 users for whom at least 50 tweets were collected.
+180 or 35.5\% of them tweet exclusively in one language (75 users tweet only in Latvian, 43 in Russian and 62 in English).
+Others tweets in several languages.
+
+To get more insight on how languages are used, we compute the language uniformity score defined as:
+\begin{equation}
+  \label{eq:score}
+  \frac{\max(n_\mathit{lv}, n_\mathit{ru}, n_\mathit{en})}{n_\mathit{lv} + n_\mathit{ru} + n_\mathit{en}}
+\end{equation}
+where $n_\mathit{lv}$ corresponds to the number of tweets in Latvian for a given user, $n_\mathit{ru}$ in Russian and $n_\mathit{en}$ in English.
+
+The higher is the score, the more dominant one language is.
+The lowest value of 0.33 means that all three languages are used equally.
+The value of 0.5 means that 50\% of tweets are written in a dominant language.
+
+The histogram on Figure~\ref{fig:score-hist} shows the score distribution.
+425 (83.8\%) users tweet mostly in one language (their scores are greater than 0.897).
+For 78  (15.4\%) users the score is between 0.5 and 0.897.
+There are only four (0.8\%) users whose dominant language share is less than 50\%.
+
+Among the four Twitter users whose score is less than 0.5---meaning that they use all three languages extensively---three are personal accounts and one is a company account.
+
+Another interesting accounts that tweet equally in Latvian and Russian, but do not tweet in English are the accounts of a library and a football club.
+It worth noting that mixing languages is not common in Latvian Internet.
+For example, large media portals tend to have separate Latvian and Russian web-sites and separate Twitter handlers.
+
+To illustrate the language usage pattern between multilingual users, their first most frequently used language, their second most frequently used language and their third frequently languages were identified.
+Latvian is not only the most used language among the monoligual users, but also is the first and second most common choice between the multilingual users.
+Russian is the most frequent third choice, refer to Table~\ref{tab:language-use}).
+
+\input{figures/language-use.tex}
+
 \bibliography{references,dmilajevs_publications}

### 2017-04-17 paper.tex
Figure position
The rest

@@
 \maketitle
+
 \begin{abstract} 
+A possibility of building a trilingual Latvian-Russian-English corpus of tweets from Riga, Latvia is investigated.
+Such a corpus, when constructed, might be of great use for training machine translation models, examining cross-lingual phenomena and studying the population of Riga.
+This pilot study shows that it is feasible to build such a resource.
+The resulted corpus is made publicaly available and might be used to construct a large comparable corpus.
 \end{abstract}
@@
 \label{sec:introduction}
+
+Comparable corpora are widely used by the natural language processing community to build machine translation or information retrieval models.
+The goal of this work is to investigate whether it is possible to build a comparable linguistic resource that originates from Riga, Latvia.
+Riga is a great location because it is a multilingual city where Latvian and Russian are widely used in everyday life, and English is spread in tourism and commerce.
+Being linked to a location, the resulting corpus might benefit not only the language engineers and linguists, but also social scientists.
+
+Twitter,\footnote{\url{https://twitter.com}} a microblog platform, is used as the data source in this pilot study.
+Over the period of 5 months (November 2016 to March 2017), tweets from Riga were collected, filtered and analysed.
+The main goal of the analysis is to investigate whether a creation of a comparable tweet corpus is feasible and what the corpus construction strategy should be.
+
+The analysis of the tweets, shows that they reflect real world events, such as national celebration or international political affair.
+While all three languages are represented---45.5\% tweets are in Latvian, 33.9\% in Russian and 20.7\% in English---not all events are equally discussed in all languages (Section~\ref{sec:timeline}).
+By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}).
+
+The properties of the corpus correspond to the expectation to reflect the real world events and language use proportion, but its size is to small to draw solid conclusions.
+However, the construction of a reliable corpus is the matter of the data collection procedure, because, as this study shows, there is relevant data on Twitter.
+
+The resulted dataset is distributed as a supplement to the paper and available online at \texttt{http://...}
 
@@
 
-The initial set of tweets was retrieved by subscribing to the \texttt{POST status/filter} endpoint of the Twitter Streaming API\footnote{\url{https://dev.twitter.com/streaming/reference/post/statuses/filter}} with Poultry \cite{dmitrijs_milajevs_2017_546609}.
+The initial set of tweets was retrieved by subscribing to the \texttt{POST status/filter} endpoint of the Twitter Streaming API\footnote{\url{https://dev.twitter.com/streaming/reference/post/statuses/filter}}.
+%with Poultry \cite{dmitrijs_milajevs_2017_546609}
 The collected tweets had to be geo-located and had to originate from the area of Riga, the capital of Latvia.\footnotemark{}
@@
 On April 14th 2017, the collection was rehydrated by querying the Twitter API with the collected tweet IDs to get rid of the deleted tweets.
-In addition, the tweets that originated retweets were included to the collection: the JSON representation of a retweet includes the original tweet, which we extracted and added to the collection.
+In addition, the tweets that originated retweets were included to the collection: the JSON representation of a retweet includes the original tweet, which was extracted and added to the collection.
 Rehydrated and expanded collection resulted to 220\,883 tweets.
@@
 
-Further analysis of the extended rehydrated collection showed that there are 23115 (10.5\%) tweets that originated from the user check-ins with Swarm\footnote{\url{https://www.swarmapp.com/}} on Foursquare.\footnote{\url{https://foursquare.com/}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' most of the time follow a predefined template and thus do not reflect real language use.
+Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.\footnote{\url{https://foursquare.com}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' most of the time follow a predefined template and thus do not reflect real language use.
 
@@
 Table~\ref{tab:client-counts} shows the top ten most popular clients in the rehydrated collection.
-Together with the tweets originated from Foursquare, tweets from Instagram, an image sharing service), and Endomondo, a workout tracking service, were removed.
-Tweets written using the ``Word Cities'' client that posts weather reports and the Linkis client, a promotion website, were also removed.
+Together with the tweets originated from Foursquare, tweets from Instagram,\footnote{\url{https://www.instagram.com}} an image sharing service, and Endomondo,\footnote{\url{https://www.endomondo.com}} a workout tracking service, were removed.
+Tweets written using the ``Word Cities'' client that posts weather reports and the Linkis client---a promotion website---were also removed.
 
-The final collection resulted into 136\,067 tweets which are in Latvian, Russian or English and created after 1st of November 2016.
+The final collection resulted into 136\,067 tweets which are in Latvian, Russian or English and created after the 1st of November 2016.
 The language of a tweet is provided by the corresponding field in the tweet JSON representation.
@@
 Out of 136\,067 tweets that constitute the final collection, 45.5\% are in Latvian, 33.9\% are in Russian and 20.7\% are in English, see Table~\ref{tab:language-counts} for tweet counts.
-The ratio between the Latvian and Russian tweets is roughly the same as the proportion of ethnic Latvians and Russians in Riga, which is 46.2\% to 37.7\%.\footnote{\url{https://en.wikipedia.org/wiki/Riga}}
+% The ratio between the Latvian and Russian tweets is roughly the same as the proportion of ethnic Latvians and Russians in Riga, which is 46.2\% to 37.7\%.\footnote{\url{https://en.wikipedia.org/wiki/Riga}}
 
@@
 Table~\ref{tab:timeline-lang-corr} suggests, that tweets in Latvian and English share similar behaviour.
-The Russian tweet timeline is distinct from both timelines, tough it's behaviour is more similar to the Latvian timeline than to English.
+The Russian tweet timeline is distinct from both timelines, tough its behaviour is more similar to the Latvian timeline than to English.
 
@@
 
+\input{figures/score-hist.tex}
+
 The inauguration of Donald Trump as the 45th President of the United States held on 20th of January 2017. The number of Latvian tweets increases, while for other languages it stays roughly the same.
@@
 
-Timeline analysis gives an insight of what events are reflected on Twitter, but does not indicate how twitter is used by ethnic communities, as language choice might depend on the topic, rather than author preference.
+Timeline analysis gives an insight of what events are reflected on Twitter, but does not indicate how Twitter is used by ethnic groups, as language choice might depend on the topic.
 
@@
 
-\input{figures/score-hist.tex}
-
 We have seen that the reflection of an event on Twitter depends on the language.
 How are languages used individually?
-Do tweet authors tweet in a one language all the time, or switch between languages depending on the context?
+Do Twitter users tweet in a one language all the time, or switch between languages depending on the context?
 
@@
 
-The higher is the score, the more dominant one language is.
-The lowest value of 0.33 means that all three languages are used equally.
+%The higher the score, the more dominant one language.
+The lowest possible value of 0.33 means that all three languages are used equally.
 The value of 0.5 means that 50\% of tweets are written in a dominant language.
+The value of 1 means that the user tweets exclusively in one language.
 
@@
 
-To illustrate the language usage pattern between multilingual users, their first most frequently used language, their second most frequently used language and their third frequently languages were identified.
+To illustrate the language usage pattern between multilingual users, their first most frequently used language, their second most frequently used language and their third most frequently language were identified.
 Latvian is not only the most used language among the monoligual users, but also is the first and second most common choice between the multilingual users.
-Russian is the most frequent third choice, refer to Table~\ref{tab:language-use}).
+Russian is the most frequent third choice, as Table~\ref{tab:language-use} shows.
 
 \input{figures/language-use.tex}
+
+\section{Conclusion}
+
+We have seen that location-based tweet collection produces adequate results. Tweets in all three target languages were collected, and the resulted collection reflects real world events.
+
+To produce reliable statistics, more data has to be collected.
+One way of getting more data (apart from a longer data collection period) is to retrieve tweets that contain words that are unique to a language \cite{sang2013}.
+This might work for Latvian, as it is reasonable to assume that a Latvian tweet originates from Latvia, but will not work for Russian and English.
+
+Instead, the properties of this corpus might be used to bootstrap data collection.
+The most active users might be identified and their tweets collected.
+If having a balanced corpus is important, then multilingual users might be preferred over monolingual.
+The user connection graph might also be useful.
+
+Regarding the social aspect of the corpus, the future work might have a look into why topics are language dependant? How separated/unified is the society? What are the controversial topics.
 

### 2017-04-18 paper.tex
Sascha changes.

@@
 \begin{abstract}
-A possibility of building a trilingual Latvian-Russian-English corpus of tweets from Riga, Latvia is investigated.
-Such a corpus, when constructed, might be of great use for training machine translation models, examining cross-lingual phenomena and studying the population of Riga.
+Twitter has become a rich source for linguistic data.
+Here, a possibility of building a trilingual Latvian-Russian-English corpus of tweets from Riga, Latvia is investigated.
+Such a corpus, once constructed, might be of great use for multiple purposes such as training machine translation models, examining cross-lingual phenomena and studying the population of Riga.
 This pilot study shows that it is feasible to build such a resource.
-The resulted corpus is made publicaly available and might be used to construct a large comparable corpus.
+The resulting corpus is made publicly available and can be used to construct a large comparable corpus.
 \end{abstract}
@@
 Comparable corpora are widely used by the natural language processing community to build machine translation or information retrieval models.
-The goal of this work is to investigate whether it is possible to build a comparable linguistic resource that originates from Riga, Latvia.
-Riga is a great location because it is a multilingual city where Latvian and Russian are widely used in everyday life, and English is spread in tourism and commerce.
-Being linked to a location, the resulting corpus might benefit not only the language engineers and linguists, but also social scientists.
+The goal of this work is to investigate whether it is possible to build a comparable linguistic resource that originates from one specific location -- Riga, Latvia.
+Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is used a lingua france in tourism and commerce.
+Being linked to a specific location, the resulting corpus might benefit not only the language engineers and linguists, but also social scientists.
 
@@
 
-The analysis of the tweets, shows that they reflect real world events, such as national celebration or international political affair.
+The analysis of the tweets, shows that they reflect real world events, such as national celebrations or international political affairs.
 While all three languages are represented---45.5\% tweets are in Latvian, 33.9\% in Russian and 20.7\% in English---not all events are equally discussed in all languages (Section~\ref{sec:timeline}).
@@
 
-The properties of the corpus correspond to the expectation to reflect the real world events and language use proportion, but its size is to small to draw solid conclusions.
+The properties of the corpus correspond to the expectation to reflect the real world events and language use proportion, but its size is too small to draw solid conclusions.
 However, the construction of a reliable corpus is the matter of the data collection procedure, because, as this study shows, there is relevant data on Twitter.
 
-The resulted dataset is distributed as a supplement to the paper and available online at \texttt{http://...}
+The resulting dataset is distributed as a supplement to the paper and available online at \texttt{http://...}
 
@@
 
-251\,083 tweets were collected from the period from the 1st of November 2016 to the 31st of March 2017.
+251\,083 tweets were collected within the period from the 1st of November 2016 to the 31st of March 2017.
 On April 14th 2017, the collection was rehydrated by querying the Twitter API with the collected tweet IDs to get rid of the deleted tweets.
-In addition, the tweets that originated retweets were included to the collection: the JSON representation of a retweet includes the original tweet, which was extracted and added to the collection.
-Rehydrated and expanded collection resulted to 220\,883 tweets.
+In addition, the tweets that originated from retweets were added to the collection: the JSON representation of a retweet includes the original tweet, which was extracted and added to the collection.
+The rehydrated and expanded collection resulted in a total of 220\,883 tweets.
 Table~\ref{tab:tweet-counts} summarises the number of collected tweets.
@@
 Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.\footnote{\url{https://foursquare.com}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' most of the time follow a predefined template and thus do not reflect real language use.
+# What does this mean and why?
 
@@
 Table~\ref{tab:client-counts} shows the top ten most popular clients in the rehydrated collection.
-Together with the tweets originated from Foursquare, tweets from Instagram,\footnote{\url{https://www.instagram.com}} an image sharing service, and Endomondo,\footnote{\url{https://www.endomondo.com}} a workout tracking service, were removed.
-Tweets written using the ``Word Cities'' client that posts weather reports and the Linkis client---a promotion website---were also removed.
+Together with the tweets originating from Foursquare, tweets from Instagram,\footnote{\url{https://www.instagram.com}} an image sharing service, and Endomondo,\footnote{\url{https://www.endomondo.com}} a workout tracking service, were removed.
+Tweets written using the ``Word Cities'' client which posts weather reports and the Linkis client---a promotion website---were also removed.
 
-The final collection resulted into 136\,067 tweets which are in Latvian, Russian or English and created after the 1st of November 2016.
+The final collection resulted in 136\,067 tweets which are in Latvian, Russian or English and created after the 1st of November 2016.
 The language of a tweet is provided by the corresponding field in the tweet JSON representation.
@@
 
-Figure~\ref{fig:timeline} shows the bandwidth of tweets in time for all three languages.
+Figure~\ref{fig:timeline} shows the bandwidth of tweets over time for all three languages.
 There are several peaks in Twitter usage.
 Some of them affect all three languages, as in early January, some of them affect only one language, as in late January.
+# Any significance?
+Ah, I see next paragraph.
 
 If the Twitter behaviour is affected by the events in the real world, then the peaks should correspond to events in the real word.
-The difference in peaks, could then be explained as there are different real word events that trigger discussions on Twitter in Latvian, Russian and English.
+The difference in peaks could then be explained as there are different real word events that trigger discussions on Twitter in Latvian, Russian and English.
 Table~\ref{tab:timeline-lang-corr} suggests, that tweets in Latvian and English share similar behaviour.
@@
 Numerous events take place across the country including candle placing on and by the wall of Riga Castle, torchlight processions and fireworks.
-The volume of tweets peaks in all three languages, thought the peak in Russian is less significant than in Latvian and English.
+The volume of tweets peaks in all three languages, though the peak in Russian is less significant than in Latvian and English.
 
@@
 
-% There is higher amount of Russian tweets, but Latvian and English tweet volume stays constant.
+% There is a higher amount of Russian tweets, but Latvian and English tweet volume stays constant.
 
@@
 How are languages used individually?
-Do Twitter users tweet in a one language all the time, or switch between languages depending on the context?
+Do Twitter users tweet in one language all the time, or switch between languages depending on the context?
 
@@
 It worth noting that mixing languages is not common in Latvian Internet.
-For example, large media portals tend to have separate Latvian and Russian web-sites and separate Twitter handlers.
+For example, large media portals tend to have separate Latvian and Russian web-sites and separate Twitter handles.
 
 To illustrate the language usage pattern between multilingual users, their first most frequently used language, their second most frequently used language and their third most frequently language were identified.
-Latvian is not only the most used language among the monoligual users, but also is the first and second most common choice between the multilingual users.
+Latvian is not only the most used language among the monoligual users but also is the first and second most common choice between the multilingual users.
 Russian is the most frequent third choice, as Table~\ref{tab:language-use} shows.
@@
 
-Regarding the social aspect of the corpus, the future work might have a look into why topics are language dependant? How separated/unified is the society? What are the controversial topics.
+Regarding the social aspect of the corpus, the future work might have a look into why topics are language dependant? How separated/unified is the society? What are the controversial topics?
 

### 2017-04-18 paper.tex
Minor

@@
 
-\input{figures/timeline.tex}
-
-The initial set of tweets was retrieved by subscribing to the \texttt{POST status/filter} endpoint of the Twitter Streaming API\footnote{\url{https://dev.twitter.com/streaming/reference/post/statuses/filter}}.
+The initial set of tweets was retrieved by subscribing to the \texttt{POST status/filter} endpoint of the Twitter Streaming API.\footnote{\url{https://dev.twitter.com/streaming/reference/post/statuses/filter}}
 %with Poultry \cite{dmitrijs_milajevs_2017_546609}
@@
 Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.\footnote{\url{https://foursquare.com}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' most of the time follow a predefined template and thus do not reflect real language use.
-# What does this mean and why?
+% What does this mean and why?
 
@@
 
-\section{Timeline analysis}
+\section{Tweet analysis}
 \label{sec:timeline}
+
+\input{figures/timeline.tex}
 
@@
 Some of them affect all three languages, as in early January, some of them affect only one language, as in late January.
-# Any significance?
+% Any significance?
 Ah, I see next paragraph.
@@
 
-\section{Language use}
+\section{User analysis}
 \label{sec:lang-use}

### 2017-04-18 paper.tex

@@
 The goal of this work is to investigate whether it is possible to build a comparable linguistic resource that originates from one specific location -- Riga, Latvia.
-Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is used a lingua france in tourism and commerce.
+Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is used a lingua franca in tourism and commerce.
 Being linked to a specific location, the resulting corpus might benefit not only the language engineers and linguists, but also social scientists.
@@
 The histogram on Figure~\ref{fig:score-hist} shows the score distribution.
-425 (83.8\%) users tweet mostly in one language (their scores are greater than 0.897).
-For 78  (15.4\%) users the score is between 0.5 and 0.897.
+420 (82.8\%) users tweet mostly in one language (their scores are greater than 0.9).
+For 83  (16.4\%) users the score is between 0.5 and 0.9.
 There are only four (0.8\%) users whose dominant language share is less than 50\%.
@@
 To illustrate the language usage pattern between multilingual users, their first most frequently used language, their second most frequently used language and their third most frequently language were identified.
-Latvian is not only the most used language among the monoligual users but also is the first and second most common choice between the multilingual users.
-Russian is the most frequent third choice, as Table~\ref{tab:language-use} shows.
+If user tweeted equally in two (three) languages, then the two (three) languages were given the maximal preference.
+A user who tweeted equally in Latvian and Russian, but less in English, will counted as Latvian and Russian being the first perference, English as the third.
+
+Latvian is not only the most used language among the monoligual users, but also is the first and third most common choice between the multilingual users.
+The preference for Russian is similar to Latvian, despite the numbers being slightly lower, suggesting its significant role in everyday life.
+English is almost the ultimate second choice, proving its role as lingua franca, as Table~\ref{tab:language-use} shows.
 

### 2017-04-20 paper.tex
Wednesday

@@
 
+% I am not sure what the following means ... are all these tweets related to real world events? I think this needs some clearity
 The analysis of the tweets, shows that they reflect real world events, such as national celebrations or international political affairs.
@@
 The resulting dataset is distributed as a supplement to the paper and available online at \texttt{http://...}
+
+\section{Related work}
+
+Twitter provides an easy way to build a large text corpora for research.
+A numerous tweet collections are build and released for a variety of purposes.
+For example, \citet{sang2013} discuss the process of building a large collection of Dutch tweets and challenges of accessing the data.
+Their retrieval method is based on a list of frequent Dutch words.
+
+\citet{SANVICENTE16.465} build a parallel multilingual corpus of tweets%
+%in Spanish, Basque, Catalan, Galician and Portuguese
+.
+Their process consists of two parts: retrieval and alignment.
+Retrieval is based on a list of multilingual users.
+The collected tweets are aligned using crowdsourcing.
+\citet{ling-EtAl:2013:ACL2013} automatically extract parallel segments from Sina Weibo (a Chinese counterpart of Twitter).
+
+\citet{gotti-langlais-farzindar:2013:LASM} use the parallel web pages mentioned in tweets of the agencies and organisations of Canada to train a statistical machine translation model.
+
+There is a body of research of the Latvian twittersphere, for example the work on sentiment analysis \cite{Peisenieks2014} and opinion mining \cite{vspats2016opinion}. Both studies focus on Latvian.
+
+% \citet{gotti-langlais-farzindar:2013:LASM,Peisenieks2014}
 

### 2017-04-20 paper.tex
Intro

@@
 Comparable corpora are widely used by the natural language processing community to build machine translation or information retrieval models.
-The goal of this work is to investigate whether it is possible to build a comparable linguistic resource that originates from one specific location -- Riga, Latvia.
+The goal of this work is to investigate whether it is possible to build a comparable linguistic resource of tweets that originates from one specific location--Riga, Latvia.
 Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is used a lingua franca in tourism and commerce.
-Being linked to a specific location, the resulting corpus might benefit not only the language engineers and linguists, but also social scientists.
 
-Twitter,\footnote{\url{https://twitter.com}} a microblog platform, is used as the data source in this pilot study.
+Despite the fact that Latvian and Russian are widely used, there is little interaction between the two ethnic communities.
+The local media consists of two subsystems (Latvian and Russian) which use different sources and present different views on current affairs \cite{muiznieks2010}.
+However, despite the fact that large media portals tend to have separate Latvian and Russian web-sites, the same opinions are found in comments to controversial content on both versions of web-sites, making the Internet a public space for a dialogue  between the ethnic communities \cite{sulmane2010}.
+The resulting corpus of user generated content is a fruitful resource for studying the integration of the two communities, by identifying what is being discussed; how, and most importantly why it is being discussed.
+
+% Twitter,\footnote{\url{https://twitter.com}} a microblog platform, is used as the data source in this pilot study.
 Over the period of 5 months (November 2016 to March 2017), tweets from Riga were collected, filtered and analysed.
@@
 
-% I am not sure what the following means ... are all these tweets related to real world events? I think this needs some clearity
-The analysis of the tweets, shows that they reflect real world events, such as national celebrations or international political affairs.
+The analysis of the tweets, shows that some real world events, such as national celebrations or international political affairs, are actively discussed on Twitter.
 While all three languages are represented---45.5\% tweets are in Latvian, 33.9\% in Russian and 20.7\% in English---not all events are equally discussed in all languages (Section~\ref{sec:timeline}).
@@
 
-Twitter provides an easy way to build a large text corpora for research.
+Twitter\footnote{\url{https://twitter.com}} provides an easy way to build a large text corpora for research.
 A numerous tweet collections are build and released for a variety of purposes.
@@
 Among the four Twitter users whose score is less than 0.5---meaning that they use all three languages extensively---three are personal accounts and one is a company account.
-
 Another interesting accounts that tweet equally in Latvian and Russian, but do not tweet in English are the accounts of a library and a football club.
-It worth noting that mixing languages is not common in Latvian Internet.
-For example, large media portals tend to have separate Latvian and Russian web-sites and separate Twitter handles.
 

### 2017-04-22 paper.tex

@@
 \usepackage{graphicx}
+\usepackage{balance}
 
 %\aclfinalcopy % Uncomment this line for the final submission
-%\def\aclpaperid{***} %  Enter the acl Paper ID here
+\def\aclpaperid{8} %  Enter the acl Paper ID here
 
@@
 
-\title{Toward a Comparable Corpora of Latvian, Russian and English Tweets}
+\title{Toward a Comparable Corpus of Latvian, Russian and English Tweets}
 
@@
 
+% No reference to Wierbicka? 
 Comparable corpora are widely used by the natural language processing community to build machine translation or information retrieval models.
@@
 
-The analysis of the tweets, shows that some real world events, such as national celebrations or international political affairs, are actively discussed on Twitter.
-While all three languages are represented---45.5\% tweets are in Latvian, 33.9\% in Russian and 20.7\% in English---not all events are equally discussed in all languages (Section~\ref{sec:timeline}).
-By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}).
+To see whether a corpus is comparable, the peaks of Twitter usage were identified.
+The peaks correspond to real world events, such as national celebrations or international political affairs, are actively discussed on Twitter in all three languages, but in different proportions (Section~\ref{sec:timeline}).
+
+All three languages are represented---45.5\% tweets are in Latvian, 33.9\% in Russian and 20.7\% in English.\footnote{
+The ratio of ethnic Latvians to Russians in Riga is 46.2\% to 37.7\%.}
+By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}), making the tweet collection by subscribing to multilingual users partial.
 
 The properties of the corpus correspond to the expectation to reflect the real world events and language use proportion, but its size is too small to draw solid conclusions.
-However, the construction of a reliable corpus is the matter of the data collection procedure, because, as this study shows, there is relevant data on Twitter.
+However, the construction of a reliable corpus is a matter of the data collection procedure, because, as this study shows, there is relevant data on Twitter.
 
-The resulting dataset is distributed as a supplement to the paper and available online at \texttt{http://...}
+The resulting dataset is available online at \texttt{http://...}
+% IMS these are all good reasons to get a corpus containing tweets from all three languages.  How can we know the corpus is comparable, and what can we do if it is?
 
@@
 
-Twitter\footnote{\url{https://twitter.com}} provides an easy way to build a large text corpora for research.
-A numerous tweet collections are build and released for a variety of purposes.
+Twitter\footnote{\url{https://twitter.com}} provides an easy way to build a large text corpus for research.
+Numerous tweet collections are built for a variety of purposes.
 For example, \citet{sang2013} discuss the process of building a large collection of Dutch tweets and challenges of accessing the data.
+%In what sense?
+Can you elaborate?
+I think this is an interesting point.
 Their retrieval method is based on a list of frequent Dutch words.
+% IMS release of twitter collections is actually fraught b/c of TOS. DM: but releasing tweets IDs is fine.
 
 \citet{SANVICENTE16.465} build a parallel multilingual corpus of tweets%
-%in Spanish, Basque, Catalan, Galician and Portuguese
+%in Spanish, Basque, Catalan, Galician and Portuguese %% All Romance languages Latvian, Russian and English are overall more diverse
 .
@@
 
-There is a body of research of the Latvian twittersphere, for example the work on sentiment analysis \cite{Peisenieks2014} and opinion mining \cite{vspats2016opinion}. Both studies focus on Latvian.
-
-% \citet{gotti-langlais-farzindar:2013:LASM,Peisenieks2014}
+There is a small but growing body of research of the Latvian Twittersphere, for example, work on sentiment analysis \cite{Peisenieks2014} and opinion mining \cite{vspats2016opinion}.
+Both studies focus on Latvian.
 
@@
 
+% IMS My old observations (2010-2012) were that very few tweets had valid geo data. Has that changed?
+
 251\,083 tweets were collected within the period from the 1st of November 2016 to the 31st of March 2017.
@@
 
-Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.\footnote{\url{https://foursquare.com}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' most of the time follow a predefined template and thus do not reflect real language use.
+Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.\footnote{\url{https://foursquare.com}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' follow a predefined template most of the time and thus do not reflect real language use.
 % What does this mean and why?
@@
 Out of 136\,067 tweets that constitute the final collection, 45.5\% are in Latvian, 33.9\% are in Russian and 20.7\% are in English, see Table~\ref{tab:language-counts} for tweet counts.
-% The ratio between the Latvian and Russian tweets is roughly the same as the proportion of ethnic Latvians and Russians in Riga, which is 46.2\% to 37.7\%.\footnote{\url{https://en.wikipedia.org/wiki/Riga}}
+% IMS the JSON field for language is actually a country tag, unless things have changed.
+Do we know anything about Twitter's language ID accuracy?
+DM: According to the Twitter docs, it's "the machine-detected language of the Tweet text".
+https://blog.twitter.com/2015/evaluating-language-identification-performance might give an answer regarding accuracy
 
@@
 
-Figure~\ref{fig:timeline} shows the bandwidth of tweets over time for all three languages.
+Figure~\ref{fig:timeline} shows the number of tweets per day over time for all three languages.
 There are several peaks in Twitter usage.
 Some of them affect all three languages, as in early January, some of them affect only one language, as in late January.
-% Any significance?
-Ah, I see next paragraph.
 
-If the Twitter behaviour is affected by the events in the real world, then the peaks should correspond to events in the real word.
+If the Twitter behaviour is affected by events in the real world, then these peaks should correspond to events in the real world.
 The difference in peaks could then be explained as there are different real word events that trigger discussions on Twitter in Latvian, Russian and English.
 Table~\ref{tab:timeline-lang-corr} suggests, that tweets in Latvian and English share similar behaviour.
-The Russian tweet timeline is distinct from both timelines, tough its behaviour is more similar to the Latvian timeline than to English.
+The Russian tweet timeline is distinct from both timelines, though its behaviour is more similar to the Latvian timeline than to the English.
+%% ADD the numbers into the text using this format?
+http://www.discoveringstatistics.com/docs/writinglabreports.pdf I would ask this from my students actually.
+I have become quite strict about reporting statistical results properly.
+See also p.
+228 here: http://www.soc.univ.kiev.ua/sites/default/files/library/elopen/andy-field-discovering-statistics-using-spss-third-edition-20091.pdf
 
 \input{figures/timeline-lang-corr.tex}
-
-What are the distinctive and similar properties of the timelines? To answer the question, we first identify the real world events that happened during the biggest peaks.
+% IMS be more specific here, you are computing the tweet frequency per unit time, right?
+What are the distinctive and similar properties of the timelines? To answer the question, we first identify the real world events that happened at the time of the highest peaks.
 
@@
 The volume of tweets peaks in all three languages, though the peak in Russian is less significant than in Latvian and English.
+% tweets peaks is close to a crash blossom. ;-)
 
@@
 There is a lot of activity during the New Year celebration for all three languages. Russian tweets peak the most almost reaching the same number of tweets as tweets in Latvian.
+% Show the sharpest rise?? I'm not sure whether peak the most is precise enough.
 
@@
 
-The inauguration of Donald Trump as the 45th President of the United States held on 20th of January 2017. The number of Latvian tweets increases, while for other languages it stays roughly the same.
+The inauguration of Donald Trump as the 45th President of the United States was held on 20th of January 2017.
+The number of Latvian tweets increases, while for other languages it stays roughly the same.
+% Surpringing, no? My German friends were posting and tweeting in English that day. This contrasts slightly with the next sentence.
 
@@
 
-We have seen that the reflection of an event on Twitter depends on the language.
+We have seen evidence of an event on Twitter reflecting an event depending on the language.
 How are languages used individually?
-Do Twitter users tweet in one language all the time, or switch between languages depending on the context?
+Do Twitter users tweet in one language all the time, or do they switch between languages depending on the context?
 
@@
 180 or 35.5\% of them tweet exclusively in one language (75 users tweet only in Latvian, 43 in Russian and 62 in English).
-Others tweets in several languages.
+Others tweet in several languages.
 
@@
 \end{equation}
-where $n_\mathit{lv}$ corresponds to the number of tweets in Latvian for a given user, $n_\mathit{ru}$ in Russian and $n_\mathit{en}$ in English.
+where $n_\mathit{lv}$ corresponds to the number of tweets in Latvian for a given user, $n_\mathit{ru}$ to the number of tweets in Russian, and $n_\mathit{en}$ to the number of tweets in English.
+% I added an Oxford comma. I assume you are typing in AmE now?
 
-%The higher the score, the more dominant one language.
+%The higher the score, the more dominant a language. %% WHY removed?
 The lowest possible value of 0.33 means that all three languages are used equally.
 The value of 0.5 means that 50\% of tweets are written in a dominant language.
+% Technically.
+it's either at 0.34 or at 0.501, no?
 The value of 1 means that the user tweets exclusively in one language.
 
-The histogram on Figure~\ref{fig:score-hist} shows the score distribution.
+The histogram in Figure~\ref{fig:score-hist} shows the score distribution.
 420 (82.8\%) users tweet mostly in one language (their scores are greater than 0.9).
@@
 Among the four Twitter users whose score is less than 0.5---meaning that they use all three languages extensively---three are personal accounts and one is a company account.
-Another interesting accounts that tweet equally in Latvian and Russian, but do not tweet in English are the accounts of a library and a football club.
+Other interesting accounts that tweet equally in Latvian and Russian, but do not tweet in English are the accounts of a library and a football club.
 
 To illustrate the language usage pattern between multilingual users, their first most frequently used language, their second most frequently used language and their third most frequently language were identified.
-If user tweeted equally in two (three) languages, then the two (three) languages were given the maximal preference.
-A user who tweeted equally in Latvian and Russian, but less in English, will counted as Latvian and Russian being the first perference, English as the third.
+If a user tweeted equally in two (three) languages, then the two (three) languages were given the maximal preference.
+A user who tweeted equally in Latvian and Russian, but less in English, will be counted as Latvian and Russian being their first preference, English as the third.
 
@@
 English is almost the ultimate second choice, proving its role as lingua franca, as Table~\ref{tab:language-use} shows.
+% as a lingua franca in this context/medium
 
@@
 
-We have seen that location-based tweet collection produces adequate results. Tweets in all three target languages were collected, and the resulted collection reflects real world events.
+We have seen that location-based tweet collection produces adequate results. Tweets in all three target languages were collected, and the resulting collection reflects real world events.
 
 To produce reliable statistics, more data has to be collected.
-One way of getting more data (apart from a longer data collection period) is to retrieve tweets that contain words that are unique to a language \cite{sang2013}.
-This might work for Latvian, as it is reasonable to assume that a Latvian tweet originates from Latvia, but will not work for Russian and English.
-
+One way of getting more data (apart from a longer data collection period) is to retrieve tweets that contain words that are unique to a language.
+This might work for Latvian, as it is reasonable to assume that a Latvian tweet originates from Latvia % If there was space for this, I would add the reference to the "1,000 languages" (book) entry which I sent you earlier as a picture via twitter messenger.
+, but this will not work for Russian and English.
 Instead, the properties of this corpus might be used to bootstrap data collection.
@@
 The user connection graph might also be useful.
-
-Regarding the social aspect of the corpus, the future work might have a look into why topics are language dependant? How separated/unified is the society? What are the controversial topics?
+Regarding the social aspect of the corpus, the future work might have a look into why topics are language dependant?
+How unified is the society?
+What are the controversial topics?
+%Sorry, just did a major change. Hope you lost nothing.
 
@@
 
+\balance
+
 \end{document}
+
+% IMS key questions I have
+% We don't know if this corpus is comparable.
+Do we care?
+How would we know?
+This is an important question because the workshop wants to know if they should take this paper.
+Will people at the workshop want to hear about this work?
+% Do the rates of laguage use in the tweets reflect the actual ethnic population or daily language usage?  If not, then our sample has bias, and maybe things we observe are due to the sample bias.
+% The discussion of the peaks is weak... it feels like we could explain away any relative curve shape.  Can we say something substantive about why only Latvian speakers tweet in late January?

### 2017-04-22 paper.tex

@@
 
-The resulting dataset is available online at \texttt{http://...}
+% The resulting dataset is available online at \texttt{http://...}
 % IMS these are all good reasons to get a corpus containing tweets from all three languages.  How can we know the corpus is comparable, and what can we do if it is?
@@
 
-Twitter\footnote{\url{https://twitter.com}} provides an easy way to build a large text corpus for research.
+Twitter %\footnote{\url{https://twitter.com}} 
+provides an easy way to build a large text corpus for research.
 Numerous tweet collections are built for a variety of purposes.
@@
 
-Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.\footnote{\url{https://foursquare.com}} This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' follow a predefined template most of the time and thus do not reflect real language use.
+Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.%\footnote{\url{https://foursquare.com}} 
+This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' follow a predefined template most of the time and thus do not reflect real language use.
 % What does this mean and why?
@@
 Table~\ref{tab:client-counts} shows the top ten most popular clients in the rehydrated collection.
-Together with the tweets originating from Foursquare, tweets from Instagram,\footnote{\url{https://www.instagram.com}} an image sharing service, and Endomondo,\footnote{\url{https://www.endomondo.com}} a workout tracking service, were removed.
-Tweets written using the ``Word Cities'' client which posts weather reports and the Linkis client---a promotion website---were also removed.
+Together with the tweets originating from Foursquare, tweets from Instagram, %\footnote{\url{https://www.instagram.com}}
+an image sharing service, and Endomondo, %\footnote{\url{https://www.endomondo.com}}
+a workout tracking service, were removed. Tweets written using the ``Word Cities'' client which posts weather reports and the Linkis client---a promotion website---were also removed.
 
@@
 
-There is a lot of activity during the New Year celebration for all three languages. Russian tweets peak the most almost reaching the same number of tweets as tweets in Latvian.
-% Show the sharpest rise?? I'm not sure whether peak the most is precise enough.
+There is a lot of activity during the New Year celebration for all three languages. Russian tweets show the sharpest rise the most almost reaching the same number of tweets as tweets in Latvian.
 
@@
 where $n_\mathit{lv}$ corresponds to the number of tweets in Latvian for a given user, $n_\mathit{ru}$ to the number of tweets in Russian, and $n_\mathit{en}$ to the number of tweets in English.
-% I added an Oxford comma. I assume you are typing in AmE now?
 
-%The higher the score, the more dominant a language. %% WHY removed?
+The higher the score, the more dominant a language.
 The lowest possible value of 0.33 means that all three languages are used equally.
 The value of 0.5 means that 50\% of tweets are written in a dominant language.
-% Technically.
-it's either at 0.34 or at 0.501, no?
 The value of 1 means that the user tweets exclusively in one language.
@@
 The preference for Russian is similar to Latvian, despite the numbers being slightly lower, suggesting its significant role in everyday life.
-English is almost the ultimate second choice, proving its role as lingua franca, as Table~\ref{tab:language-use} shows.
-% as a lingua franca in this context/medium
+English is almost the ultimate second choice, proving its role as a lingua franca, as Table~\ref{tab:language-use} shows.
 
@@
 One way of getting more data (apart from a longer data collection period) is to retrieve tweets that contain words that are unique to a language.
-This might work for Latvian, as it is reasonable to assume that a Latvian tweet originates from Latvia % If there was space for this, I would add the reference to the "1,000 languages" (book) entry which I sent you earlier as a picture via twitter messenger.
-, but this will not work for Russian and English.
+This might work for Latvian, as it is reasonable to assume that a Latvian tweet originates from Latvia, % If there was space for this, I would add the reference to the "1,000 languages" (book) entry which I sent you earlier as a picture via twitter messenger.
+but this will not work for Russian and English.
 Instead, the properties of this corpus might be used to bootstrap data collection.
@@
 What are the controversial topics?
-%Sorry, just did a major change. Hope you lost nothing.
 

### 2017-04-26 paper.tex
Updates from ShareLaTeX

@@
 The goal of this work is to investigate whether it is possible to build a comparable linguistic resource of tweets that originates from one specific location--Riga, Latvia.
-Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is used a lingua franca in tourism and commerce.
+Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is used as a lingua franca in tourism and commerce.
 
@@
 To see whether a corpus is comparable, the peaks of Twitter usage were identified.
-The peaks correspond to real world events, such as national celebrations or international political affairs, are actively discussed on Twitter in all three languages, but in different proportions (Section~\ref{sec:timeline}).
+The peaks correspond to real world events, such as national celebrations, international political affairs or weather are actively discussed on Twitter in all three languages, but in different proportions (Section~\ref{sec:timeline}).
 
@@
 The ratio of ethnic Latvians to Russians in Riga is 46.2\% to 37.7\%.}
-By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}), making the tweet collection by subscribing to multilingual users partial.
+By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}), making the tweet collection by subscribing only to multilingual users inclomplete.
 
 The properties of the corpus correspond to the expectation to reflect the real world events and language use proportion, but its size is too small to draw solid conclusions.
-However, the construction of a reliable corpus is a matter of the data collection procedure, because, as this study shows, there is relevant data on Twitter.
+However, the construction of a reliable comparable corpus is a matter of the data collection procedure and corpus' application, because, as this study shows, not all topics are discussed equally.
 
@@
 \citet{ling-EtAl:2013:ACL2013} automatically extract parallel segments from Sina Weibo (a Chinese counterpart of Twitter).
-
+%
 \citet{gotti-langlais-farzindar:2013:LASM} use the parallel web pages mentioned in tweets of the agencies and organisations of Canada to train a statistical machine translation model.
@@
 The rehydrated and expanded collection resulted in a total of 220\,883 tweets.
-Table~\ref{tab:tweet-counts} summarises the number of collected tweets.
+%Table~\ref{tab:tweet-counts} summarises the number of collected tweets.
 
-\input{figures/tweet-counts.tex}
+% \input{figures/tweet-counts.tex}
 
-Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare.%\footnote{\url{https://foursquare.com}} 
+Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare. %\footnote{\url{https://foursquare.com}} 
 This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' follow a predefined template most of the time and thus do not reflect real language use.
@@
 an image sharing service, and Endomondo, %\footnote{\url{https://www.endomondo.com}}
-a workout tracking service, were removed. Tweets written using the ``Word Cities'' client which posts weather reports and the Linkis client---a promotion website---were also removed.
+a workout tracking service, were removed. Tweets written using the ``Word Cities'' client, which posts weather reports, and the Linkis client---a promotion website---were also removed.
 
@@
 18 November is the Proclamation Day of the Republic of Latvia.
-Numerous events take place across the country including candle placing on and by the wall of Riga Castle, torchlight processions and fireworks.
-The volume of tweets peaks in all three languages, though the peak in Russian is less significant than in Latvian and English.
-% tweets peaks is close to a crash blossom. ;-)
+Also, on the 8th of November, the
+United States presidential election took place.
+
+The number of tweets significantly increased for Latvian and English, and not so much for Russian.
+Manual inspection of the tweets in that period reveals that the US elections are discussed in all three languages, while the national celebrations of the 11th and the 18th of November are mostly discussed in Latvian.
+The discussion includes such topics as the news related to celebrations, historical notes, reminders of working hours of businesses, greeting and advertisement.
+
+Manual inspection also shows that events are language sensitive.
+For example, the election results were discussed by Latvians in English.
+Also, businesses reported their working hours during the national celebrations in Latvian and do not duplicate this information in Russian.
 
@@
 
-There is a lot of activity during the New Year celebration for all three languages. Russian tweets show the sharpest rise the most almost reaching the same number of tweets as tweets in Latvian.
+In early January a snowstorm hit Riga.
+In Latvian and Russian the discussed topics were the same, namely, appreciation of snow, the transportation difficulties and outdoor photos.
+Tweets in English mostly contained photos showing how beautiful Latvia is in Winter.
 
@@
 
-The inauguration of Donald Trump as the 45th President of the United States was held on 20th of January 2017.
+The inauguration of the 45th President of the United States was held on 20th of January 2017.
 The number of Latvian tweets increases, while for other languages it stays roughly the same.
+The reason why there are relatively little politics-oriented Russian tweets might be that 60\% of citizens and 47\% of non-citizens are interested in politics \cite{civicsociety}.
+Out of citizens, 60\% are ethnic Latvians, 27\% are ethnic Russians.
+Out of non-citizens, 66\% are Russians, and less than 1\% are Latvians.
+% https://lv.wikipedia.org/wiki/Nepilso%C5%86i_(Latvija)
 % Surpringing, no? My German friends were posting and tweeting in English that day. This contrasts slightly with the next sentence.
-
-% \paragraph{Early March}
-
-% There are peaks in Latvian and Russian tweets.
-
-% \paragraph{Mid March}
-
-% There is a general decline in Twitter usage for Latvian and English, but not for Russian, when for a short period of time there are more Russian tweets posted than Latvian.
-
-Timeline analysis gives an insight of what events are reflected on Twitter, but does not indicate how Twitter is used by ethnic groups, as language choice might depend on the topic.
 
@@
 
-We have seen evidence of an event on Twitter reflecting an event depending on the language.
-How are languages used individually?
-Do Twitter users tweet in one language all the time, or do they switch between languages depending on the context?
+We have seen an evidence that topics are languge dependant. How many Twitter users switch between languages?
 
@@
 
-To produce reliable statistics, more data has to be collected.
+How comparable is the corpora?
+Topics are language dependent, so it is not the case that all topics are discussed in every language.
+There are ``monolingual topics'' such as the independence day in Latvia.
+Even ``multilingual topics'' vary in content, as with the snowstorm tweets, where Latvian and Russian tweets shared common topics, but tweets in English were distinct.
+
+%
+The final answer is that it depends on the application.
+For machine translation, it is important to have similar content, so the parallel segments can be extracted, for example from Latvian and Russian snowstorm tweets.
+For a social study, the corpus has to be representative, so that the topics could lead to the analysis of the communities as why the president inauguration was discussed much less in Russian than in Latvian.
+
+%To produce reliable statistics, more data has to be collected.
 One way of getting more data (apart from a longer data collection period) is to retrieve tweets that contain words that are unique to a language.
 This might work for Latvian, as it is reasonable to assume that a Latvian tweet originates from Latvia, % If there was space for this, I would add the reference to the "1,000 languages" (book) entry which I sent you earlier as a picture via twitter messenger.
-but this will not work for Russian and English.
+%but this will not work for Russian and English.
 Instead, the properties of this corpus might be used to bootstrap data collection.
 The most active users might be identified and their tweets collected.
-If having a balanced corpus is important, then multilingual users might be preferred over monolingual.
+If having a balanced corpus is important, then multilingual users might be preferred over monolingual, bun in case of Riga, there are not so many of them.
 The user connection graph might also be useful.
-Regarding the social aspect of the corpus, the future work might have a look into why topics are language dependant?
-How unified is the society?
-What are the controversial topics?
+
+%Regarding the social aspect of the corpus, the future work might have a look into why topics are language dependant? How unified is the society? What are the controversial topics?
+
+\balance
 
@@
 \bibliographystyle{acl_natbib}
-
-\balance
 

### 2017-04-27 paper.tex

@@
 an image sharing service, and Endomondo, %\footnote{\url{https://www.endomondo.com}}
-a workout tracking service, were removed. Tweets written using the ``Word Cities'' client, which posts weather reports, and the Linkis client---a promotion website---were also removed.
+a workout tracking service, were removed. Tweets written using the World Cities client, which posts weather reports, and the Linkis client---a promotion website---were also removed.
 
@@
 
-How comparable is the corpora?
+How comparable are the language samples within the corpus?
 Topics are language dependent, so it is not the case that all topics are discussed in every language.
@@
 For machine translation, it is important to have similar content, so the parallel segments can be extracted, for example from Latvian and Russian snowstorm tweets.
-For a social study, the corpus has to be representative, so that the topics could lead to the analysis of the communities as why the president inauguration was discussed much less in Russian than in Latvian.
+For a social study, the corpus has to be representative, so that the topics can lead to the analysis of the communities as in the case of why the president inauguration was discussed much less in Russian than in Latvian.
 

### 2017-05-22 paper.tex
Updates from ShareLaTeX

@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
+\aclfinalcopy % Uncomment this line for the final submission
 \def\aclpaperid{8} %  Enter the acl Paper ID here
@@
 
-\author{First Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\}
+\author{Dmitrijs Milajevs \\
+  NIST \\
+  Gaithersburg, MD, USA \\
+  {\tt dmitrijs.milajevs@nist.gov}}
 
@@
 Such a corpus, once constructed, might be of great use for multiple purposes such as training machine translation models, examining cross-lingual phenomena and studying the population of Riga.
-This pilot study shows that it is feasible to build such a resource.
-The resulting corpus is made publicly available and can be used to construct a large comparable corpus.
+This pilot study shows that it is feasible to build such a resource by building and analysing a pilot corpus, which is made publicly available and can be used to construct a large comparable corpus.
 \end{abstract}
@@
 Comparable corpora are widely used by the natural language processing community to build machine translation or information retrieval models.
-The goal of this work is to investigate whether it is possible to build a comparable linguistic resource of tweets that originates from one specific location--Riga, Latvia.
+The goal of this work is to investigate in  apilot study whether it is possible to build a comparable linguistic resource of tweets that originates from one specific location--Riga, Latvia.
 Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is used as a lingua franca in tourism and commerce.
@@
 
-Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins with on Foursquare. %\footnote{\url{https://foursquare.com}} 
+Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins on Foursquare. %\footnote{\url{https://foursquare.com}} 
 This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' follow a predefined template most of the time and thus do not reflect real language use.
@@
 Out of citizens, 60\% are ethnic Latvians, 27\% are ethnic Russians.
-Out of non-citizens, 66\% are Russians, and less than 1\% are Latvians.
+Out of non-citizens, 66\% are Russians, and less than 1\% are Latvians.\footnotemark{}
+
+\footnotetext{\href{https://lv.wikipedia.org/wiki/Nepilso\%C5\%86i_(Latvija)}{\tt{https://lv.wikipedia.org/wiki/ Nepilso\c{n}i\_(Latvija)}}}
 % https://lv.wikipedia.org/wiki/Nepilso%C5%86i_(Latvija)
@@
 English is almost the ultimate second choice, proving its role as a lingua franca, as Table~\ref{tab:language-use} shows.
-
+%
 \input{figures/language-use.tex}

### 2017-05-22 paper.tex
Updates from ShareLaTeX

@@
 However, despite the fact that large media portals tend to have separate Latvian and Russian web-sites, the same opinions are found in comments to controversial content on both versions of web-sites, making the Internet a public space for a dialogue  between the ethnic communities \cite{sulmane2010}.
-The resulting corpus of user generated content is a fruitful resource for studying the integration of the two communities, by identifying what is being discussed; how, and most importantly why it is being discussed.
+A corpus of user generated content is a fruitful resource for studying the integration of the two communities, by identifying what is being discussed; how, and most importantly why it is being discussed.
 

### 2017-05-22 paper.tex
Update

@@
 However, despite the fact that large media portals tend to have separate Latvian and Russian web-sites, the same opinions are found in comments to controversial content on both versions of web-sites, making the Internet a public space for a dialogue  between the ethnic communities \cite{sulmane2010}.
-A corpus of user generated content is a fruitful resource for studying the integration of the two communities, by identifying what is being discussed; how, and most importantly why it is being discussed.
+A corpus of user generated content that comes from one location is a fruitful resource for studying the integration of the two communities, by identifying what is being discussed; how, and most importantly why it is being discussed.
 
 % Twitter,\footnote{\url{https://twitter.com}} a microblog platform, is used as the data source in this pilot study.
-Over the period of 5 months (November 2016 to March 2017), tweets from Riga were collected, filtered and analysed.
-The main goal of the analysis is to investigate whether a creation of a comparable tweet corpus is feasible and what the corpus construction strategy should be.
 
-To see whether a corpus is comparable, the peaks of Twitter usage were identified.
+Over the period of 5 months (November 2016 to March 2017), tweets from Riga were collected, filtered and analysed.\footnotemark{} The main goal of the analysis is to investigate whether a creation of a comparable tweet corpus is feasible and what the corpus construction strategy should be.
+To see whether the corpus is comparable, the peaks of Twitter usage were identified.
 The peaks correspond to real world events, such as national celebrations, international political affairs or weather are actively discussed on Twitter in all three languages, but in different proportions (Section~\ref{sec:timeline}).
+
+\footnotetext{\url{https://doi.org/10.5281/zenodo.582300}}
 
@@
 The ratio of ethnic Latvians to Russians in Riga is 46.2\% to 37.7\%.}
-By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}), making the tweet collection by subscribing only to multilingual users inclomplete.
+By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}), making the tweet collection by subscribing only to multilingual users incomplete.
 

### 2017-05-22 paper.tex
Intro

@@
 Here, a possibility of building a trilingual Latvian-Russian-English corpus of tweets from Riga, Latvia is investigated.
-Such a corpus, once constructed, might be of great use for multiple purposes such as training machine translation models, examining cross-lingual phenomena and studying the population of Riga.
-This pilot study shows that it is feasible to build such a resource by building and analysing a pilot corpus, which is made publicly available and can be used to construct a large comparable corpus.
+Such a corpus, once constructed, might be of great use for multiple purposes including  training machine translation models, examining cross-lingual phenomena and studying the population of Riga.
+This pilot study shows that it is feasible to build such a resource by collecting and analysing a pilot corpus, which is made publicly available and can be used to construct a large comparable corpus.
 \end{abstract}
@@
 Comparable corpora are widely used by the natural language processing community to build machine translation or information retrieval models.
-The goal of this work is to investigate in  apilot study whether it is possible to build a comparable linguistic resource of tweets that originates from one specific location--Riga, Latvia.
-Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is used as a lingua franca in tourism and commerce.
+The goal of this work is to investigate in  a pilot study whether it is possible to build a comparable linguistic resource of tweets that originates from one specific location--Riga, Latvia.
+Riga is a great location for this because it is a multilingual city in which Latvian and Russian are both widely used in everyday life, and English is a lingua franca in tourism and commerce.
 
@@
 The local media consists of two subsystems (Latvian and Russian) which use different sources and present different views on current affairs \cite{muiznieks2010}.
-However, despite the fact that large media portals tend to have separate Latvian and Russian web-sites, the same opinions are found in comments to controversial content on both versions of web-sites, making the Internet a public space for a dialogue  between the ethnic communities \cite{sulmane2010}.
-A corpus of user generated content that comes from one location is a fruitful resource for studying the integration of the two communities, by identifying what is being discussed; how, and most importantly why it is being discussed.
+Even though large media portals tend to have separate Latvian and Russian web-sites, the same opinions are found in comments to controversial content on both versions of web-sites, making the Internet a public space for a dialogue  between the ethnic communities \cite{sulmane2010}.
+A corpus of user generated content from Riga is a fruitful resource for studying the integration of the two communities, by identifying what is being discussed; how, and most importantly why it is being discussed.
 
-% Twitter,\footnote{\url{https://twitter.com}} a microblog platform, is used as the data source in this pilot study.
-
-Over the period of 5 months (November 2016 to March 2017), tweets from Riga were collected, filtered and analysed.\footnotemark{} The main goal of the analysis is to investigate whether a creation of a comparable tweet corpus is feasible and what the corpus construction strategy should be.
-To see whether the corpus is comparable, the peaks of Twitter usage were identified.
-The peaks correspond to real world events, such as national celebrations, international political affairs or weather are actively discussed on Twitter in all three languages, but in different proportions (Section~\ref{sec:timeline}).
+The pilot corpus\footnotemark{} consists of tweets over the period of 5 months (November 2016 to March 2017).
+The main goal of the analysis is to investigate whether a creation of a comparable tweet corpus is feasible and what the corpus construction strategy should be.
+To see whether the pilot corpus is comparable, the peaks of Twitter usage were identified.
+These peaks correspond to real world events (national celebrations, international political affairs and weather).
+The events are actively discussed in all languages, but in different proportions (Section~\ref{sec:timeline}).
 
@@
 
-All three languages are represented---45.5\% tweets are in Latvian, 33.9\% in Russian and 20.7\% in English.\footnote{
+All three languages are represented: 45.5\% tweets are in Latvian, 33.9\% in Russian and 20.7\% in English.\footnote{
 The ratio of ethnic Latvians to Russians in Riga is 46.2\% to 37.7\%.}
-By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}), making the tweet collection by subscribing only to multilingual users incomplete.
+By studying users' tweeting habits, we see that the majority of users (83.3\%) mostly tweets in one language (Section~\ref{sec:lang-use}), making the tweet collection strategy that considers only multilingual users incomplete.
 
@@
 
+\section*{Acknowledgements}
+
+The author thanks Dr.~Sascha Griffiths for valuable discussion and Tatiana Chepurko for literature suggestions.
+
+\section*{Disclaimer}
+
+Certain commercial products are identified in this paper in order to specify the experimental procedure adequately.
+Such identification is not intended to imply recommendation or endorsement by the National Institute of Standards and Technology, nor is it intended to imply that the identified products are necessarily the best available for the purpose.
+
 \balance

### 2017-05-22 paper.tex
Minor

@@
 
-The properties of the corpus correspond to the expectation to reflect the real world events and language use proportion, but its size is too small to draw solid conclusions.
+The properties of the corpus correspond to the expectation that it will reflect the real world events and language use proportion, but its size is too small to draw solid conclusions.
 However, the construction of a reliable comparable corpus is a matter of the data collection procedure and corpus' application, because, as this study shows, not all topics are discussed equally.
@@
 If a user tweeted equally in two (three) languages, then the two (three) languages were given the maximal preference.
-A user who tweeted equally in Latvian and Russian, but less in English, will be counted as Latvian and Russian being their first preference, English as the third.
+A user who tweeted equally in Latvian and Russian, but less in English, is counted as Latvian and Russian being their first preference, English as the third.
 

### 2017-05-26 paper.tex
Final

@@
 The main goal of the analysis is to investigate whether a creation of a comparable tweet corpus is feasible and what the corpus construction strategy should be.
-To see whether the pilot corpus is comparable, the peaks of Twitter usage were identified.
+To see whether the pilot corpus is comparable, the peaks of Twitter usage were analysed.
 These peaks correspond to real world events (national celebrations, international political affairs and weather).
@@
 
-% IMS My old observations (2010-2012) were that very few tweets had valid geo data. Has that changed?
-
 251\,083 tweets were collected within the period from the 1st of November 2016 to the 31st of March 2017.
-On April 14th 2017, the collection was rehydrated by querying the Twitter API with the collected tweet IDs to get rid of the deleted tweets.
-In addition, the tweets that originated from retweets were added to the collection: the JSON representation of a retweet includes the original tweet, which was extracted and added to the collection.
+On April 14th 2017, the collection was rehydrated\footnotemark{} by querying the Twitter API with the collected tweet IDs to get rid of the deleted tweets.
+In addition, the tweets that
+%
+\footnotetext{Since distribution of raw tweet data is not allowed, tweets IDs are shared instead. Hydration is the process of retrieval of raw tweet data by IDs.}
+%
+originated from retweets were added to the collection: the JSON\footnotemark{} representation of a retweet includes the original tweet, which was extracted and added to the collection.
 The rehydrated and expanded collection resulted in a total of 220\,883 tweets.
@@
 
+\footnotetext{\url{http://json.org}}
+
 % \input{figures/tweet-counts.tex}
 
-Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins on Foursquare. %\footnote{\url{https://foursquare.com}} 
+Further analysis of the extended rehydrated collection showed that there are 23\,115 (10.5\%) tweets that originated from check-ins on Foursquare.\footnote{\url{https://foursquare.com}}
 This motivated additional filtering of the rehydrated collection, as ``check-in tweets'' follow a predefined template most of the time and thus do not reflect real language use.
